FraCaS: Temporal Analysis

### 2020-06-23 paper2020/paper.tex
revisions: background

\paragraph{GF}
The grammatical framework (GF, \citet{ranta_grammatical_2004}) is a powerful parser
generator for natural languages, based on type-theoretical abstract grammars.
This work relies a syntactic representation of the FraCaS test suite in GF
abstract syntax~\citep{Ljunglof:2012}. Thus, we skip the parsing phase
and avoid any syntactic ambiguity.

For the purpose of this paper, the important point is that GF the
syntax aims at a balance of sufficient abstraction to provide a
semantically-relevant structure, but at the same time it embeds
sufficiently many syntactic features to be able to reconstruct
unambiguously natural-language text. \cref{ex:nobel}
shows an example abstract syntax tree and its realisation in English.

\paragraph{Coq}
Coq is an interactive theorem prover (proof assistant). It is based on
the calculus of coinductive constructions~\citep{werner_une_1994} We
do not use any coinduction (or even induction) in this paper, relying
on the pure lambda-calculus inner core of Coq.  Coq is a very
powerful reasoning engine that makes it fit for implementing natural
language semantics in it but also, on a more practical side, for the
the task of NLI, when the latter is formalized as a theorem proving
task. Coq supports, as already mentioned, dependent typing, but also
subtyping, which are instrumental in expressing NL semantics.

@@

\paragraph{Dynamic Semantics}
We additionally assume support for (non-temporal) anaphora, using a
monadic-based dynamic semantics
\cite{unger_dynamic_2011,charlow_monadic_2015,charlow_modular_2017}.
Our particular implementation is not perfect \ifanon (citations
anonymized) \else \citet{bernardy_wide-coverage_2019} \fi but non-temporal
anaphoroi in the testsuite are generally resolved as they should be.

As for other basic phenomena, there is not much interaction between
our treatment of time and non-temporal anaphora. One critical
exception is discussed in \cref{sec:repeatable-achievements}.

@@

In this paper we borrow several notions from classical temporal
semantics such as ``stative'', ``achievement'', ``activity'', etc.,
even though our definitions do not perfectly match the classical
ones. We explain our precise meaning forthese terms in the body of the
paper. Nevertheless, we refer the reader to \citet{steedman:2000} for
an extensive review of temporal semantics.

For the \textit{cognoscenti}, we can already point out some
differences in terminology: we use the term activity as a general term
which encompasses the three classical notions of activites,
achievements and accomplishments. Indeed, insofar as the test suite is
concerned, we find that these three categories can be collapsed into a
single one.  As such, we think it is a good classification, given that
they generally afford the correct inferences for the testsuite. (In
this paper, we always assume that the FraCaS testsuite is correct.)

@@

\begin{multline*}
\constant{unicity}_P : P (t_1,t_2) → P (t_3,t_4) → \\ (t_1 = t_3) ∧ (t_2 = t_4)
\end{multline*}

>>>>

\paragraph{GF}
The grammatical framework (GF, \citet{ranta_grammatical_2004}) is a powerful parser
generator for natural languages, based on type-theoretical abstract grammars.
The present work leverages a syntactic representation of the FraCaS test suite in GF
abstract syntax~\citep{Ljunglof:2012}. Thanks to this, we skip the parsing phase
and avoid any syntactic ambiguity.

For the purpose of this paper, the important point is that the GF
syntax aims at a balance of sufficient abstraction to provide a
semantically-relevant structure, but at the same time it embeds
sufficiently many syntactic features to be able to reconstruct
natural-language text. That is, the parse trees generally satisfy the
homomorphism requirement of \citet{Montague:EFL70,Montague1974}, and
we can focus on the translation of syntactic trees to logical forms.
\cref{ex:nobel} shows an example abstract syntax tree and its
realisation in English.

\paragraph{Coq}
Coq is an interactive theorem prover (proof assistant). It is based on
the calculus of coinductive constructions~\citep{werner_une_1994} We
do not use any coinduction (or even induction) in this paper, relying
on the pure lambda-calculus inner core of Coq.  Coq is a very powerful
reasoning engine that makes it fit for implementing natural language
semantics.  Coq supports, as already mentioned, dependent typing, but
also subtyping. Both concepts are instrumental in expressing NL
semantics\citet{TODO: chatzikyriakidis and luo}.  Besides, on a more
practical side, it works well for the the task of NLI, when the latter
is formalized as a theorem proving task: its many tactics mean that
many tasks in theorem proving are trivialized. In particular, all
problems of time-intervals inclusion, which occur in every temporal
problems, are solved with Coq's linear arithmetic tactic.

@@

\paragraph{Dynamic Semantics}
We additionally assume support for (non-temporal) anaphora, using a
monadic-based dynamic semantics, generally following the state of the
art in this matter
\citep{unger_dynamic_2011,charlow_monadic_2015,charlow_modular_2017}.
Our particular implementation does not handle every case \ifanon
(citations anonymized) \else \citet{bernardy_wide-coverage_2019} \fi
but non-temporal anaphoroi in the testsuite are generally resolved as
they should be: on the whole accuracy is not affected significantly by
issues in this subsystem.

As it is the case for other basic phenomena, there is not much
interaction between our treatment of time and non-temporal
anaphora. One critical exception is discussed in
\cref{sec:repeatable-achievements}.

@@

In this paper we borrow several notions from classical temporal
semantics such as ``stative'', ``achievement'', ``activity'', etc.,
even though our definitions do not perfectly match the classical
ones. We explain our precise meaning for these terms in the body of the
paper. Nevertheless, we refer the reader to \citet{steedman:2000} for
an extensive review of formal temporal semantics.

For the \textit{cognoscenti}, we can already point out some
differences in terminology: we use the term activity as a general term
which encompasses the three classical notions of activites,
achievements and accomplishments. Indeed, insofar as the test suite is
concerned, we find that these three categories can be collapsed into a
single one (they are subject to \cref{eq:unicity}).  As such, we think
it is a good classification, given that it generally affords the
correct inferences for the testsuite. (In this paper, we always assume
that the problems in the FraCaS testsuite are correctly annotated.)

@@

\begin{multline*}
  \constant{unicity}_P : P (t_1,t_2) → P (t_3,t_4) → \\ (t_1 = t_3) ∧ (t_2 = t_4)
  \label{eq:unicity}
\end{multline*}
